Supplement to Example Manuscript Template for a Data Analysis Project

### 2024-08-06 products/manuscript/Manuscript.qmd
update supplementary information

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

-   https://www.google.com/publicdata/directory
-   https://www.data.gov/
-   https://github.com/awesomedata/awesome-public-datasets
-   https://registry.opendata.aws/
-   https://www.kaggle.com/datasets

- https://www.kaggle.com/datasets/jainaru/electric-vehicle-population

# Summary/Abstract
_Write a summary of your project._

The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL). It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.

{{< pagebreak >}}

# Introduction

**- (a) Briefly describe what the data is, how it was collected, where you will get (or got) it from. How many observations do you have, what was measured? Anything else important to report about the data?**

The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL). It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.

**- (b) At this stage you are not required to already have and show the data, but if you do, even better. Then add a few lines of code which load the data and using some of the commands you learned about, provide summary descriptions of the data.**
```{r}
data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
data <- read.csv(data_location)
str(data)
head(data)
summary(data)
colSums(is.na(data))
```

**- (c) Explain the question you want to answer using the data. What will be your outcome(s) of interest (if any)? What (if any) specific predictors will you focus on? What relations/patterns are you looking for in the data?**

Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape ?

## Objectives:

- Understand Geographic Distribution: Determine the distribution of electric vehicles across different counties and cities in Washington State.
- Identify Popular Models: Identify the most commonly registered electric vehicle makes and models.
- Analyze Electric Range: Investigate the distribution of electric range among registered electric vehicles.-
- Explore Clean Alternative Fuel Vehicle Eligibility: Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.

## Outcomes of Interest
- Geographic Distribution: The distribution of electric vehicles across counties and cities in Washington State.
- Popular Models: The most frequently registered electric vehicle makes and models.
- Electric Range Analysis: The range distribution and average electric range of registered electric vehicles.
- Clean Alternative Fuel Vehicle Eligibility: The proportion of vehicles eligible for clean alternative fuel vehicle incentives.

## Predictors
- Geographic Variables: County, City, Postal Code.
- Vehicle Attributes: Model Year, Make, Model, Electric Vehicle Type, Electric Range.

## Relations/Patterns
- Geographic Distribution: Analyzing clusters of electric vehicles in specific regions.
- Popular Models: Identifying dominant electric vehicle makes and models.
- Electric Range Analysis: Understanding the distribution of electric ranges and potential implications for consumer preferences.
- Clean Alternative Fuel Vehicle Eligibility: Examining eligibility patterns and potential factors influencing eligibility.
- Correlations: Investigating any associations between electric vehicle characteristics and legislative districts or census tracts, which may offer insights into policy implications and consumer behavior.

**(d) As much as you know, suggest how you will analyze it. At this stage in the course, we haven’t covered analysis approaches yet, so you can keep things vague and non-technical here.**

- Descriptive Analysis: Summarize the dataset to understand its contents, including summary statistics and frequency tables.

-Visualization: Use plots and charts to visualize aspects like geographic distribution, electric range, and vehicle types.

- Comparative Analysis: Compare different groups within the dataset, such as BEVs vs PHEVs or eligibility for incentives by vehicle type.

- Correlation Analysis: Explore potential relationships between variables, such as electric range and base MSRP.

- Geospatial Analysis (optional): Use R packages to analyze spatial patterns in the data.

- Exploratory Data Analysis (EDA): Engage in exploratory analysis to identify interesting patterns or trends.

**(e) You are allowed, but not yet required, to provide background information for the question you plan to answer. For instance you can describe why it’s an interesting question, who else has done similar analyses, how your analysis will be new/different, etc. Similar to what you read in an introduction to a research paper. For the final report, you’ll need these parts. For part 1, they are not required, but you are welcome to already write down some of that. **

## Primary Research Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape in Washington State?

Sub-questions:

-  How are electric vehicles distributed across different counties, cities, and States?
-  Which EV makes and models are most popular?
-  What is the range of electric vehicles in terms of electric range and base MSRP?
-  How does eligibility for clean alternative fuel vehicle incentives vary among different vehicle types?
-  Are there any significant correlations between EV characteristics and geographic or legislative factors?

**(d) Eventually, for your final report, what you write for this part will go into different sections of the full report. Some will go into the introduction, some in the methods section. You can already place these items there, or for now just write them as a single section.**

**Introduction**
Electric vehicles (EVs), including Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs), are crucial in the global effort to reduce greenhouse gas emissions and dependence on fossil fuels. This dataset, sourced from the Washington State Department of Licensing (DOL) and obtained from Kaggle, provides a comprehensive overview of registered electric vehicles in Washington State. Analyzing this data offers valuable insights into the adoption patterns and characteristics of EVs, informing policies and infrastructure development to further promote sustainable transportation.

**Objectives:**

**Geographic Distribution:** Determine the distribution of electric vehicles across different counties and cities.
**Popular Models:** Identify the most commonly registered EV makes and models.
**Electric Range Analysis:** Investigate the distribution of electric range among registered EVs.
**CAFV Eligibility:** Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.
**Correlation Analysis:** Explore potential correlations between EV characteristics and geographic or legislative factors.

**Methods**

**Data Cleaning and Preparation:**

- Address missing values and ensure correct data types.
- Transform variables as necessary for analysis.

To ensure the dataset is ready for analysis, the following steps were undertaken:

**1. Data Import and Initial Inspection:**

- The dataset Electric_Vehicle_Population_Data.csv was imported.
- Summary statistics and data types were displayed to understand the initial structure and content of the data.
- UPDATE (6.17.2024) - Downloaded IRS tax data to join with the EV data by zip code
  - Tax data provides many financial variables that we will include in our models (i.e. can we predict the amount of EVs per zip code based off tax data?)
- Added a geojson file to draw zip codes on a map, will use for visualizations but still figuring out how to elegantly filter the json

**2. Checking for Duplicates:**

- Identified and counted the number of duplicate rows in the dataset to avoid redundant information.
- Result: num_duplicates duplicates were found.

**3. Counting Rows with Zero MSRP:**

- Counted the number of rows where the Base.MSRP (Manufacturer's Suggested Retail Price) is zero to identify incomplete pricing data.
- Result: count_zero_msrp rows with zero MSRP were identified.

**4. Checking for Null Values:**

- Calculated the total number of null (NA) values in the dataset to address missing data.
- Result: total_na null values were found.

**5. Removing Null Values and Identifying Outliers:**

- Filtered out rows with null values to ensure data completeness.
- Identified the maximum Base.MSRP for each electric vehicle type to spot potential outliers.
- Result:The maximum MSRP for each vehicle type was summarized.

**6. Filtering Out Outliers and Zero MSRP:**

Removed the row with an extreme outlier Base.MSRP of 845,000 to prevent distortion in the analysis.
Also filtered out rows where Base.MSRP is zero to ensure only valid pricing data is considered.
By performing these steps, the dataset was cleaned and prepared, ensuring it is more accurate and reliable for subsequent analysis.

**Descriptive Statistics:**

- Compute summary statistics (mean, median, mode, standard deviation) for numerical variables.
- Generate frequency tables for categorical variables.

**Boxplot - Price Distribution vs. Electric Vehicle Type:**

This boxplot compares the price distribution of Battery Electric Vehicles (BEV) and Plug-in Hybrid Electric Vehicles (PHEV).
It shows that BEVs have a wider price range with a higher median price compared to PHEVs, which have a more compact price distribution with some outliers.

**Boxplot - Price Distribution vs. Electric Vehicle Type (Clean Alternative Fuel Vehicle Eligible):**

This boxplot illustrates the price distribution based on eligibility for Clean Alternative Fuel Vehicle incentives.
Vehicles eligible for the incentive tend to have a higher price range compared to those that are not eligible due to low battery range.

**Boxplot - Price Distribution vs. Electric Vehicle Type (Drive Range):**

This boxplot shows the distribution of drive range for clean alternative fuel vehicle eligible and non-eligible vehicles.
Eligible vehicles have a wider and higher range compared to non-eligible vehicles, which have a limited drive range.

**Scatterplot - Electric Vehicle Range Growth over Time:**

This scatterplot depicts the growth in electric vehicle range over time.
There is a clear trend showing that the electric range of vehicles has generally increased over the years.

**Boxplot - Price Distribution by Brand:**

This boxplot compares the price distribution of electric vehicles by different brands.
Tesla has the highest price range, followed by Porsche. Brands like KIA and Chrysler have lower price distributions.

**Scatterplot - Price vs. Range Correlation:**

This scatterplot shows the correlation between the vehicle price and its electric range.
There is a positive correlation where higher-priced vehicles tend to have a longer electric range.
These plots provide insights into the price and range variations of electric vehicles based on type, eligibility for incentives, brand, and over time.

**Data Visualization:**

- Create bar charts, histograms, and pie charts to visualize distributions.
- Use scatter plots and box plots to explore relationships between variables.
- Generate geographic maps to visualize EV distribution across the state.

**Comparative Analysis:**

- Compare BEVs and PHEVs in terms of electric range and base MSRP.
- Analyze differences in CAFV eligibility across makes and models.

**Correlation Analysis:**

- Calculate correlation coefficients for numerical variables.
- Use cross-tabulations to explore relationships between categorical and numerical variables.

**Geospatial Analysis (if applicable):**

- Use geographic mapping tools to identify clusters or patterns in vehicle registrations.

**Exploratory Data Analysis (EDA):**

-Perform in-depth analysis to identify trends, patterns, or outliers.
-Use advanced visualization techniques for deeper insights.

**Statistical Model**
- First iteration of models
- Produced a few statistically significant tax metrics for predicting the number of EVs in a given county
- Washington demographics datasets could further enhance our analysis

**Conclusion:**

By analyzing the electric vehicle registration data for Washington State, this project aims to uncover significant patterns and trends that can inform policy decisions and infrastructure planning. The insights gained from this analysis will contribute to the broader understanding of EV adoption and support the transition towards a more sustainable transportation future.

{{< pagebreak >}}

>>>>

---
title: "Using IRS Data to Predict Electric Vehicle Count"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

## Abstract
This study investigates the use of advanced machine learning models to predict vehicle count by zip code, utilizing a dataset enriched with socio-economic, demographic, and regional features. The primary aim is to accurately estimate vehicle counts, aiding in efficient resource distribution and urban planning. We implemented and evaluated several models, including Decision Tree, Random Forest, XGBoost, and HistGradientBoosting, focusing on their accuracy, interpretability, and computational efficiency.

Each model was trained and validated using cross-validation and a hold-out test set, with performance metrics such as Mean Absolute Error (MAE), Mean Squared Error (MSE), and R-squared used to assess their effectiveness. Among the models, the Random Forest model demonstrated superior performance, offering the best balance of predictive accuracy and model robustness. It effectively captured the complex interactions between the features and the target variable, outperforming the other models in terms of both accuracy and generalization.

The results highlight the Random Forest model's effectiveness in handling diverse and complex datasets, making it a valuable tool for predicting vehicle counts in different zip codes. This study provides critical insights into the practical applications of machine learning in urban planning and resource management, particularly in the context of transportation and infrastructure development.

## Introduction
The increasing adoption of electric vehicles (EVs) is a pivotal shift in the transportation sector, driven by growing environmental concerns, advancements in technology, and supportive government policies. Understanding the factors that influence the distribution of EVs is essential for developing targeted policies and infrastructure, such as charging stations and incentives. However, EV adoption is not uniform across regions; it varies significantly based on socio-economic and demographic factors. This study aims to explore these variations by predicting vehicle counts, particularly EVs, at the zip code level using a comprehensive dataset that combines electric vehicle registration data with individual income tax statistics from the Internal Revenue Service (IRS).

The IRS data offers detailed insights into various financial attributes of taxpayers by zip code, including income levels, deductions, and credits. By integrating this data with EV registration data, we can analyze the relationship between financial capacity, tax-related benefits, and vehicle ownership patterns. Such an analysis is crucial for identifying the socio-economic determinants of EV adoption, which can inform policymakers and urban planners.

The main objectives of this study are twofold: first, to develop predictive models that accurately estimate vehicle counts by zip code, and second, to identify and quantify the impact of key socio-economic factors on these counts. We employed a variety of machine learning models, including Decision Tree, Random Forest, XGBoost, and HistGradientBoosting, to achieve these objectives. These models were chosen for their ability to handle complex, structured data and provide interpretable results.

The findings of this study are expected to contribute valuable insights into the socio-economic factors driving EV adoption. By identifying the key predictors of vehicle counts, we can better understand the barriers to and facilitators of EV adoption in different communities. This knowledge is crucial for designing effective policies and interventions that promote sustainable transportation solutions. Furthermore, the study's methodology, which includes rigorous data cleaning, feature transformation, and model validation, provides a robust framework for future research in this area.

## Methods
In this study, we employed a variety of machine learning models to predict vehicle counts by zip code using socio-economic and demographic features extracted from IRS tax data. The data preparation, feature transformation, and model training processes are detailed below.

#### Data Preparation
The dataset was compiled by merging electric vehicle registration data with IRS tax data, focusing specifically on zip codes within the state of Washington. This geographical focus provided a diverse socio-economic landscape, enhancing the robustness of our analysis. The dataset included features such as 'Net premium tax credit amount,' 'Home mortgage interest paid amount,' 'Real estate taxes amount,' and several others related to income, tax credits, and deductions.

To prepare the dataset for analysis, we first addressed data quality issues. We standardized column names using regular expressions, ensuring consistency across the merged dataset. This step was essential for preventing errors during data processing and analysis. We also applied data filtering to include only the relevant zip codes from Washington state, further refining our dataset for localized analysis.

#### Feature Transformation
Given that many of the predictor variables exhibited skewed distributions, we applied a log transformation to normalize these features. This transformation was particularly necessary for financial features, such as 'Salaries and wages amount' and 'Home mortgage interest paid amount,' which spanned several orders of magnitude. The transformation formula used was log⁡(x+1)log(x+1), which effectively handled zero values and stabilized variance across the dataset. Normalizing the data in this manner helped improve the performance and interpretability of our machine learning models by reducing the impact of outliers and facilitating a more uniform distribution of feature values.

#### Model Development
We implemented four different machine learning models: Decision Tree, Random Forest, XGBoost, and HistGradientBoosting. Each model was selected for its unique strengths in handling structured data and complex relationships between features.

#### Model Training and Validation
Each model was trained on the processed dataset, with training and validation sets created using cross-validation techniques. Cross-validation helped ensure that the models generalize well to unseen data, preventing overfitting. We evaluated model performance using metrics such as Root Mean Squared Error (RMSE) and R-squared, which provided insights into the accuracy and explanatory power of each model.

Overall, the methods employed in this study—from data preparation and feature transformation to model development and validation—were designed to rigorously test and compare the predictive capabilities of multiple machine learning models. These methods provided a comprehensive framework for analyzing the socio-economic factors influencing vehicle counts by zip code, with a particular emphasis on the adoption of electric vehicles.

## Results
This section presents the comparative analysis of various machine learning models used to predict vehicle count by zip code. The models evaluated include XGBoost, Decision Tree, HistGradientBoosting, and Random Forest. The primary evaluation metrics considered were Root Mean Squared Error (RMSE) and R-squared, which provide insights into the accuracy and explanatory power of the models.

### Model Performance Comparison
#### XGBoost
  - RMSE: 71.27 <br/>
  - R-squared: 0.91 <br/>
  - XGBoost demonstrated strong predictive capabilities, achieving a relatively low RMSE of 71.27. The model's R-squared value of 0.91 indicates that it explains 91% of the variance in the vehicle count data, suggesting good model fit and predictive accuracy. <br/>
![XGBoost](../../results/output/model_results/XGBoost_Regressor.png)

#### Decision Tree
  - RMSE: 100.77<br/>
  - R-squared: 0.83<br/>
  -  The Decision Tree model had the highest RMSE among the evaluated models at 100.77, reflecting less precise predictions. Its R-squared value of 0.83 indicates that it explains 83% of the variance, making it the least accurate model in this comparison.
![Decision Tree](../../results/output/model_results/Decision_Tree_Regressor.png)

#### HistGradientBoosting
  -  RMSE: 83.42<br/>
  -  R-squared: 0.88<br/>
  -  The HistGradientBoosting model offered a moderate performance with an RMSE of 83.42 and an R-squared value of 0.88. While better than the Decision Tree model, it still lagged behind XGBoost and Random Forest in terms of prediction accuracy and variance explanation.
![HistGradientBoosting](../../results/output/model_results/HistGradientBoosting_Regressor.png)

#### Random Forest
  -  RMSE: 62.16<br/>
  -  R-squared: 0.935<br/>
  -  The Random Forest model outperformed all other models, achieving the lowest RMSE of 62.16 and the highest R-squared value of 0.935. These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 93.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
![Random Forest](../../results/output/model_results/Random_Forest_Regressor.png)

### Best Model Selection

Based on the comparative analysis of the models, the Random Forest model emerged as the best performing model. It achieved the lowest RMSE and the highest R-squared, demonstrating superior accuracy and generalization capability compared to the other models. The Random Forest model's ensemble nature allows it to effectively capture complex patterns in the data, reduce overfitting through bootstrapping, and handle interactions between predictor variables more adeptly. These attributes collectively contributed to its superior performance, making it the most suitable choice for predicting vehicle counts by zip code in this study.

The excellent performance of the Random Forest model suggests it is well-suited for applications requiring high prediction accuracy and robustness, particularly in urban planning and resource allocation scenarios where accurate vehicle count predictions are crucial.

## Discussion
The objective of this study was to develop machine learning models capable of predicting vehicle counts by zip code, using various socio-economic and demographic features derived from IRS tax data. The analysis identified the Random Forest model as the most accurate and robust among the evaluated models, including XGBoost, Decision Tree, and HistGradientBoosting. In addition to the model performance, a detailed examination of the top features contributing to the model predictions provides valuable insights into the socio-economic factors influencing vehicle counts, particularly for electric vehicles (EVs).

### Key Feature Analysis

The top 10 features contributing to the model predictions include:

    Net Premium Tax Credit Amount: This feature represents the tax credit amount received by individuals under the Affordable Care Act. A higher tax credit may indicate lower income levels, suggesting a potential relationship between income and the ability to purchase vehicles, including EVs.

    Home Mortgage Interest Paid Amount: This feature reflects the interest paid on home mortgages, often correlating with homeownership and wealth. Areas with higher mortgage interest payments may have a higher concentration of middle to upper-income households, which can influence vehicle ownership patterns, including the adoption of EVs.

    Real Estate Taxes Amount: Real estate taxes are indicative of property values and, indirectly, wealth. Higher real estate taxes often correspond to higher property values, which may correlate with a higher prevalence of EV ownership due to greater financial capacity.

    Self-Employed Health Insurance Deduction Amount: This deduction is available to self-employed individuals who pay their health insurance premiums. It reflects the presence of self-employed individuals in a region, who may have different financial and vehicle ownership patterns compared to salaried employees.

    Salaries and Wages Amount: This feature directly indicates income levels from employment. Higher aggregate salaries and wages in a zip code suggest greater disposable income, which can facilitate the purchase of vehicles, including newer and potentially more expensive EVs.

    Limited State and Local Taxes: This feature represents the capped amount of state and local taxes that can be deducted from federal taxes. The deduction cap, especially in high-tax states, can impact disposable income and financial planning, potentially influencing vehicle purchasing decisions.

    Home Mortgage from Personal Seller Amount: This feature indicates transactions involving home mortgages from personal sellers. While less common, these transactions can indicate unique financial arrangements in certain areas, which might correlate with specific economic profiles affecting vehicle ownership.

    Individual Retirement Arrangement (IRA) Payments Amount: Contributions to IRAs reflect financial planning and the presence of disposable income. Areas with higher IRA payments may have wealthier populations, potentially increasing the likelihood of EV adoption due to greater financial resources.

    Self-Employed (Keogh) Retirement Plans Amount: Similar to IRAs, Keogh plans are retirement savings plans for self-employed individuals. Contributions to these plans suggest higher income and financial planning, which can correlate with higher rates of vehicle ownership, including EVs.

    Additional Medicare Tax Amount: This tax applies to high-income earners and indicates a population segment with substantial earnings. Higher additional Medicare tax payments correlate with higher income levels, potentially leading to increased EV adoption due to greater financial capacity.

### Implications and Socio-Economic Insights

The analysis of these top features reveals a consistent pattern: higher vehicle counts, particularly for EVs, are associated with higher income levels, wealth indicators (such as mortgage interest and real estate taxes), and financial planning (such as retirement contributions). These factors suggest that regions with greater financial capacity are more likely to adopt EVs, reflecting broader trends in consumer behavior and market accessibility.

The emphasis on features related to self-employment and additional taxes for high earners highlights the impact of income diversity on vehicle ownership patterns. The presence of deductions and credits for self-employed individuals suggests that entrepreneurial regions may exhibit unique vehicle ownership trends, possibly favoring flexibility and lower-cost vehicles or, conversely, reflecting higher disposable income that supports EV adoption.

## Conclusion
The Random Forest model's ability to incorporate and weigh these diverse socio-economic features underscores its robustness and suitability for this type of predictive analysis. The insights gained from the top features not only enhance our understanding of the factors driving vehicle ownership but also offer valuable guidance for policymakers and urban planners aiming to promote EV adoption. Future research could expand on these findings by incorporating additional data sources, such as environmental factors or public transportation availability, to provide a more comprehensive view of the determinants of vehicle ownership and EV adoption.

{{< pagebreak >}}
## Citations
Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/><br/>

Internal Revenue Service. (n.d.). SOI tax stats - Individual income tax statistics - ZIP code data (SOI). Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi

### 2024-08-06 products/manuscript/supplement/Supplementary-Material.qmd
update supplementary information

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
#bibliography: ../../../assets/dataanalysis-references.bib
csl: ../../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

@@

A quick overview of what readers can find in the supplement.

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

{{< pagebreak >}}

@@

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","tables","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

>>>>

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
# bibliography: ../../assets/dataanalysis-references.bib
# csl: ../../assets/apa.csl
---

@@

An overview of the correct process to generate the final outputs as well as some exploratory analysis.

# Code Information

**Run analysis in the following order.**

1. Pre-Processing File (Quarto): <br/>
/code/processing-code/pre_process_tax_data.qmd - preprocesses the tax/ev data and merges them into one dataset.

2. Exploratory Data Analysis (Quarto): <br/>
/code/eda-code/eda-ev-dataset.qmd - exploratory analysis of the electric vehicle dataset <br/>
/code/eda-code/eda-ev-tax.qmd - exploratory analysis of the tax dataset <br/>

3. Predictive models (Quarto): <br/>
/code/analysis-code/statistical_analysis.qmd - builds multiple predictive models, visualizes, and compares them.

@@

#### Decision Tree Regressor:
A basic model that splits the data into subsets based on feature values, creating a tree structure. We set hyperparameters such as the maximum depth and minimum samples per leaf to control the model's complexity and prevent overfitting.

#### Random Forest Regressor:
An ensemble learning method that constructs multiple decision trees and aggregates their outputs to improve prediction accuracy. We configured key parameters like the number of trees (n_estimators) and the maximum depth, optimizing for a balance between bias and variance.

#### XGBoost Regressor:
A gradient boosting model known for its efficiency and performance. XGBoost builds trees sequentially, with each new tree aiming to correct errors made by the previous ones. We fine-tuned parameters such as the learning rate (eta), maximum depth, and regularization terms (lambda and alpha) to enhance model performance.

#### HistGradientBoosting Regressor:
A histogram-based implementation of gradient boosting that is particularly efficient for large datasets. This model discretizes continuous features into bins, reducing memory usage and speeding up training. We set parameters including the number of iterations (max_iter), learning rate, and number of bins (max_bins) to optimize the model.

# Additional results
## Log-Transformed Histogram Examples
![SE Health Insurance Deduction](../../../results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)

![Salary & Wages Amount](../../../results/figures/histograms/Salaries%20and%20wages%20amount_histogram.png)

![Net Premium Tax Credit Amount](../../../results/figures/histograms/Net%20premium%20tax%20credit%20amount_histogram.png)

## Log-Transformed Boxplot Examples
![SE Health Insurance Deduction](../../../results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)

![Salary & Wages Amount](../../../results/figures/boxplots/Salaries%20and%20wages%20amount_boxplot.png)

![Net Premium Tax Credit Amount](../../../results/figures/boxplots/Net%20premium%20tax%20credit%20amount_boxplot.png)

## Model Comparison - MAE
!['MAE'](../../../results/output/model_comparison/MAE.png)

## Model Comparison - MSE
!['MSE'](../../../results/output/model_comparison/MSE.png)

## Model Comparison - R-squared
!['R-squared'](../../../results/output/model_comparison/R-squared.png)

## Model Comparison - RMSE
!['RMSE'](../../../results/output/model_comparison/RMSE.png)

# References
- Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/>

- Internal Revenue Service. (n.d.). SOI tax stats - Individual income tax statistics - ZIP code data (SOI). Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi
